Leadership across cultures and contexts: an exploratory analysis

### 2020-05-01 leadership_across_cultures_contexts.Rmd
Minor changes in discussion.

ZG then created a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially directly interact with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders. ZG and KS then coded all text records for group structure.

@@

Leadership in residential subgroups (e.g., hunting party, work group) was relatively common among hunter-gatherers but relatively rare among agriculturists. Leadership at the supracommunity level, in contrast, was relatively rare among hunter-gatherers but quite common among agriculturists. The only example of supracommunity leadership among hunter-gatherers was from the Andamanese, a group of populations of Negrito hunter-gatherers of the Andaman Islands by [@az02-001, p. 45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section).

@@

Relationships between leadership and high status, including dependencies and causality are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019]. Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in tribal warfare or social movements). cheng_dominance_2019 and cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways. Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans social status may be a necessary but insufficient condition for achieving and maintaing positions of leadership. In our view, social context is critical to these discussions, as is often mentioned but also under-emphasized and not well supported empirically. Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses. The "low status" male who initiates a inter-group raid on a neighboring group may not be widely respected within his community nor be a contending Big man, however, he might be of exceptionally high status among his peers and high ranking within what @garfield_review_2019 [p. 71] called the "intragenerational social hierarchy."

>>>>

ZG then created a measure of group structure coded as: *residential subgroup*, which includes informal groups of co-residents, social groups, age-based groups, performance groups; *kin group*, which includes lineages, phraties, and other kin-based organizations; *economic group*, which includes subsistence groups, market groups, and other groups with primarily economic goals; *military group*, which includes all groups related to inter-group conflict; *religious group*, which includes groups collected for spiritual or supernatural purpose; *political group (community)*, which includes political groups at the level of the community, i.e., the political leader can potentially interact directly with most community members such as most village headmen, or *political group (supracommunity)*, which includes political groups that encompass multiple residential communities, such as complex chiefdoms, regional political leaders, and kings or state-level leaders. ZG and KS then coded all text records for group structure.

@@

Leadership in residential subgroups (e.g., social groups, performance groups) was relatively common among hunter-gatherers but relatively rare among agriculturists. Leadership at the supracommunity level, in contrast, was relatively rare among hunter-gatherers but quite common among agriculturists. The only example of supracommunity leadership among hunter-gatherers was from the Andamanese, a group of populations of Negrito hunter-gatherers of the Andaman Islands by [@az02-001, p. 45], who discussed the roles of a "head chief" from a permanent residential location and their marginal influence over other, local residential leaders (see quote in Methods section).

@@

Relationships between leadership and high status, including dependencies and causality are puzzling and currently the focus of much work in evolutionary leadership studies [e.g., @van_vugt_dual_2019; @smith_leadership_2020; @cheng_why_2020; @cheng_dominance_2019; @von_rueden_dynamics_2019]. Drawing on a comparative-evolutionary perspective, incorporating findings from nonhuman social species, @van_vugt_dual_2019 emphasize that although leadership and high social status do indeed often covary, leaders are not necessarily high-status individuals, exemplified when low-ranking individuals in nonhuman species emerge as leaders of group movement or when younger and lower status men and women initiate cooperative activities (such as raids in tribal warfare or social movements). cheng_dominance_2019 and cheng_why_2020, emphasize, however, that social hierarchies form principally in response to inter-individual variation in status, which may be gained via distinct pathways. Thus, according to @cheng_why_2020, although all leaders are not necessarily high status, especially across nonhuman species, among humans high social status may be a necessary but insufficient condition for achieving and maintaing positions of leadership. In our view, social context is critical to these discussions, as is often mentioned but also under-emphasized and not well supported empirically. Processes of leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social statuses. The "low status" male who initiates a inter-group raid on a neighboring group may not be widely respected within his community nor be a contending Big man, however, he might be of exceptionally high status among his peers and high ranking within what @garfield_review_2019 [p. 71] called the "intragenerational social hierarchy."
